---
title: "FINAL PROJECT"
author: "Victor Zhang"
date: "2023-03-06"
output: pdf_document
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = F,
                      message = F, 
                      warning = F) 

```

```{r}

library(tidyverse)
library(readr)
library(knitr)
library(zoo)
library(dplyr)
library(lubridate)
library(fpp2)
library(tsibble)
library(seasonal)
library(seasonalview)
library(xts)

```


# Executive Summary





# Introduction

```{r}

```

















# Data Analysis and Experimental Design


```{r}

Flex_CPI <- read_csv(file = "C:/Users/victo/Desktop/project/COREFLEXCPIM159SFRBATL.csv")

Unemployment_Rate <- read_csv(file = "C:/Users/victo/Desktop/project/UNRATE.csv")

Federal_Effective_Rate <- read_csv(file = "C:/Users/victo/Desktop/project/DFF.csv")

T10Y_2Y  <- read_csv(file = "C:/Users/victo/Desktop/project/T10Y2Y.csv")


```



# Time Series Modelling and Forecasting


```{r}

```



# Discussion and Conclusions

```{r}

```























\newpage

# Bibliography and Appendix


```{r}

```



Board of Governors of the Federal Reserve System (US)., 2023. *Federal Funds Effective Rate [DFF]*. retrieved from FRED, Federal Reserve Bank of St. Louis., [Online], Accessed:  6 March 2023 \
Available:  https://fred.stlouisfed.org/series/DFF, March 6, 2023.



Board of Governors of the Federal Reserve System., 2023. *Monetary Policy*. [Online], Accessed:  1 March 2023 \
Available: https://www.federalreserve.gov/monetarypolicy.htm



Bryan, Michael F., and Brent Meyer., 2010. “Are Some Prices in the CPI More Forward Looking Than Others? We Think So”. Federal Reserve Bank of Cleveland, Economic Commentary., [Online], \
Accessed:  1 March 2023 \
Available: https://doi.org/10.26509/frbc-ec-201002



Chatfield C., 2003. *The analysis of time series: an introduction*. CHAPMAN & HALL/CRC.



Federal Reserve Bank of New York., 2007. *Federal funds*. Fedpoints.



Federal Reserve Bank of St. Louis,. 2023. *10-Year Treasury Constant Maturity Minus 2-Year Treasury Constant Maturity [T10Y2Y]*. retrieved from FRED, Federal Reserve Bank of St. Louis., [Online], \
Accessed:  6 March 2023 \
Available: https://fred.stlouisfed.org/series/T10Y2Y



Federal Reserve Bank of Atlanta., 2023. *Flexible Price Consumer Price Index less Food and Energy [COREFLEXCPIM159SFRBATL]*. retrieved from FRED, Federal Reserve Bank of St. Louis., \
[Online], Accessed:  6 March 2023 \
Available: https://fred.stlouisfed.org/series/COREFLEXCPIM159SFRBATL



Kenny, Thomas., 2022. “The U.S. Treasury Yield Spread”. The Balance., [Online], \
Accessed:  1 March 2023 \
Available: https://www.thebalancemoney.com/2-to-10-year-us-treasury-yield-spread-1976-present-417139



Peters, G.W., 2023. *Lecture Notes*. UCSB.



Picardo, Elvis,. 2022. “How the Unemployment Rate Affects Everybody”. Investopedia., [Online], \
Accessed:  1 March 2023  \
Available: https://www.investopedia.com/articles/economics/10/unemployment-rate-get-real.asp



Rob J Hyndman and George Athanasopoulos., 2018. *Forecasting: Principles and Practice*. OTexts.



U.S. Bureau of Labor Statistics., 2023. *Unemployment Rate [UNRATE]*. retrieved from FRED, Federal Reserve Bank of St. Louis., [Online], Accessed:  6 March 2023 \
Available: https://fred.stlouisfed.org/series/UNRATE











